Revisiting the hydrodynamic modulation of short surface waves by longer waves

### 2025-03-13 paper.tex
Revision 1 (#1)
* Figures to quantify surface velocity error

* Compute wave action conservation criteria

* Homogeneity and stationarity from analytical and numerical solutions

* WIP wave action conservation limits

* WIP homogeneity and stationarity figures

* WIP appendices

* Add ref to Clamond & Dutykh

* Update velocity error plot

* Add section on wave action limits

* Appendix A

* Fix typo

* Update plotting scripts to reflect updated functions in 2wave

* Fix x-label and contour levels

* Fixes to figures

* Total derivative for acceleration in diagram

* Lagrangian gravity

* Updat

To quantify the contribution of the nonlinear terms in (\ref{eq:wavenumber})
and (\ref{eq:wave_action2}), and to evaluate the steadiness assumption of the
analytical solutions, we now proceed to numerically integrate the full set of
crest and action conservation equations.
(\ref{eq:wavenumber}) and (\ref{eq:wave_action2}) are discretised using
second-order central finite difference in space and integrated in time using the
fourth-order Runge-Kutta method \citep{butcher1996runge}.
The space is divided into 128 grid points.
This is effectively the same equation set and numerical configuration as that of
\citet{peureux2021unsteady}, except that their long-wave orbital velocities are
evaluated at $z=0$ instead of $z=\eta_L$, thus neglecting the Stokes drift
induced by long waves on short-wave groups
\citep{stokes1847,van2018stokes,monismith2020stokes}.
Their gravity modulation also does not consider the nonlinear, Lagrangian, or
curvature effects, however, this difference does not qualitatively affect the results.
Another difference is the choice of the numerical scheme for spatial differences,
which in their case is the more sophisticated MUSCL4 scheme \citep{kurganov2000new}.
Although the second-order central finite difference is not appropriate for many
numerical problems, in our case it is sufficient because it is conservative and
the fields that we compute derivatives of are smooth and continuous.
The maximum relative error of the centred finite difference in this model
is $\approx 0.066\%$.
The wavenumber and wave action are conservative over time within
$\mathcal{O}(10^{-7})$ and $\mathcal{O}(10^{-5})$ relative error, respectively.
We show in the next subsection that our numerical solutions for infinite long-wave
trains are qualitatively equivalent to those of \citet{peureux2021unsteady}.

>>>>

To quantify the contribution of the nonlinear terms in (\ref{eq:wavenumber})
and (\ref{eq:wave_action2}), and to evaluate the steadiness assumption of the
analytical solutions, we now proceed to numerically integrate the full set of
crest and action conservation equations.
(\ref{eq:wavenumber}) and (\ref{eq:wave_action2}) are discretised using
second-order central finite difference in space and integrated in time using the
fourth-order Runge-Kutta method \citep{butcher1996runge}.
The space is divided into 128 grid points.
This is effectively the same equation set and numerical configuration as that of
\citet{peureux2021unsteady}, except that their long-wave orbital velocities are
evaluated at $z=0$ instead of $z=\eta_L$, thus neglecting the Stokes drift
induced by long waves on short-wave groups
\citep{stokes1847,van2018stokes,monismith2020stokes}.
Their gravity modulation also does not consider the nonlinear, Lagrangian, or
slope effects, however, this difference does not qualitatively affect the results.
Another difference is the choice of the numerical scheme for spatial differences,
which in their case is the more sophisticated MUSCL4 scheme \citep{kurganov2000new}.
Although the second-order central finite difference is not appropriate for many
numerical problems, in our case it is sufficient because it is conservative and
the fields that we compute derivatives of are smooth and continuous.
The maximum relative error of the centred finite difference in this model
is $\approx 0.066\%$.
The wavenumber and wave action are conservative over time within
$\mathcal{O}(10^{-7})$ and $\mathcal{O}(10^{-5})$ relative error, respectively.
We show in the next subsection that our numerical solutions for infinite long-wave
trains are qualitatively equivalent to those of \citet{peureux2021unsteady}.
